Linear vertical shear instability in astrophysical disks: a numerical study

### 2014-11-30 linear_iso.tex
extensions

\section{Vertical shear instability of locally
  isothermal perturbations}
The linear problem simplifies considerably for the case
$\gamma=1$. That is, for locally isothermal perturbations where
\begin{align}
  W=Q,
\end{align}
so that
\begin{align}
  \delta v_z = \frac{1}{\ii\sigma}\frac{dQ}{dz}. 
\end{align}
Inserting this into the expression for $\delta v_x$ and the continuity
equation, we arrive at a single equation for the linear stability
problem,
\begin{align}\label{iso_ode}
  \frac{d^2Q}{dz^2} + \left(\frac{d\ln{\rho}}{dz} - \frac{\ii k_x
      r}{D}\frac{d\Omega^2}{dz}\right) \frac{dQ}{dz} +
  \sigma^2\left(\frac{1}{c_s^2} + \frac{k_x^2}{D}\right)Q=0, 
\end{align} 
subject to appropriate boundary conditions.

@@

Linearizing the integral Eq. \ref{integral_relation} and taking the
imaginary part, we have
\begin{align}
  \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
  \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
  \He_n^2(\hat{z}) d\hat{z} \notag\\
  = \delta q h \hat{k} 
  \int_{-\infty}^{\infty}
  w(\zhat)\hat{z}\He_n(\hat{z})\frac{d\He_n}{d\zhat} d\hat{z}
\end{align}
Recognizing $\hat{z} = \He_1$ and using $d\He_n/d\hat{z} = n
\He_{n-1}$ we have
 \begin{align}
   \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
   \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
   \He_n^2(\hat{z}) d\hat{z} \notag\\
   =\delta q h \hat{k} 
   \int_{-\infty}^{\infty}
   w(\zhat)\He_1(\hat{z})\He_n(\hat{z})n\He_{n-1}(\hat{z})d\hat{z}. 
 \end{align}
% Integrating the right hand side by parts and recognizing $\He_2 =
% \hat{z}^2 -1$, we have
% \begin{align}
%   \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
%   \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
%   \He_n^2(\hat{z}) d\hat{z} \notag\\
%   = \frac{1}{2}\delta q h \hat{k} 
%   \int_{-\infty}^{\infty}
%   w(\zhat)\He_2(\hat{z})\He_n^2(\hat{z})d\hat{z}. 
% \end{align}
Finally, we evaluate the integrals using following properties of Hermite polynomials
\begin{align}
  \int_{-\infty}^{\infty}
  w(\xi) \He_k(\xi)\He_l(\xi) d\xi = \sqrt{2\pi}k!\delta_{kl}, 
\end{align}
and
\begin{align}
  \int_{-\infty}^{\infty}
  w(\xi) \He_m(\xi)\He_k(\xi)\He_l(\xi) d\xi =
  \frac{\sqrt{2\pi}m!k!l!}{(s-m)!(s-k)!(s-l)!}, 
\end{align}
where $s = (m+k+l)/2$. We then obtain
% \begin{align}
%  \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
%  \left(1+\hat{k}^2\right) = \delta q h \hat{k} \frac{n!}{(n-1)!}.
% \end{align
 \begin{align}
  \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
  \left(1+\hat{k}^2\right) = \delta q h \hat{k} n.
 \end{align}
Inserting the original eigenfrequency $\sigma = \pm
\sqrt{n}\Omega_k(1+\hat{k}^2)^{-1/2}$ gives
\begin{align}
  \frac{\imag(\delta\sigma)}{\Omega_k} = \pm \frac{1}{2}\delta q h
  \sqrt{n} \frac{\hat{k}}{\sqrt{1+\hat{k}^2}}. 
\end{align}
This result agrees with Eq. \ref{simple_growth}.  
%for $n=1$, since the
%function $\He_1$ solves the governing equation exactly. 
For $\hat{k}\gg 1$ we have
\begin{align}
  \frac{\imag(\delta\sigma)}{\Omega_k} \simeq \pm \frac{1}{2}\delta q h
  \sqrt{n} \sgn{\hat{k}}. 
\end{align}

>>>>

\section{Vertical shear instability of locally
  isothermal perturbations}
{\bf need to update notation, use $\epsilon$ instead of $h$}
The linear problem simplifies considerably for the case
$\Gamma=\gamma=1$. Then 
\begin{align}
  W=Q. 
\end{align}
%so that
%\begin{align}
 % \delta v_z = \frac{1}{\ii\sigma}\frac{dQ}{dz}. 
%\end{align}
%Inserting this into the expression for $\delta v_x$ and the continuity
%equation, 
Inserting this into Eq. \ref{ode_w} gives a single equation for the linear stability
problem for isothermal perturbations in a vertically isothermal disk, 
\begin{align}\label{iso_ode}
  \frac{d^2Q}{dz^2} + \left(\frac{d\ln{\rho}}{dz} - \frac{\ii k_x
      r}{D}\frac{d\Omega^2}{dz}\right) \frac{dQ}{dz} +
  \sigma^2\left(\frac{1}{c_s^2} + \frac{k_x^2}{D}\right)Q=0, 
\end{align} 
subject to appropriate boundary conditions.

We remark that Eq .\ref{iso_ode} is also applicable for
$\gamma=\Gamma\neq 1$ in the limit $t_c\to\infty$, since in that case
$W\simeq Q$ as well. The following discussion therefore also holds for
vertically non-isothermal, adiabatic disks.

@@

Linearizing the integral Eq. \ref{integral_relation} and taking the
imaginary part, we have
\begin{align}
  \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
  \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
  \He_n^2(\hat{z}) d\hat{z} \notag\\
  = \delta q h \hat{k} 
  \int_{-\infty}^{\infty}
  w(\zhat)\hat{z}\He_n(\hat{z})\frac{d\He_n}{d\zhat} d\hat{z}
\end{align}
Recognizing $\hat{z} = \He_1$ and using $d\He_n/d\hat{z} = n
\He_{n-1}$ we have
 \begin{align}
   \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
   \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
   \He_n^2(\hat{z}) d\hat{z} \notag\\
   =\delta q h \hat{k} 
   \int_{-\infty}^{\infty}
   w(\zhat)\He_1(\hat{z})\He_n(\hat{z})n\He_{n-1}(\hat{z})d\hat{z}. 
 \end{align}
% Integrating the right hand side by parts and recognizing $\He_2 =
% \hat{z}^2 -1$, we have
% \begin{align}
%   \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
%   \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
%   \He_n^2(\hat{z}) d\hat{z} \notag\\
%   = \frac{1}{2}\delta q h \hat{k} 
%   \int_{-\infty}^{\infty}
%   w(\zhat)\He_2(\hat{z})\He_n^2(\hat{z})d\hat{z}. 
% \end{align}
Finally, we evaluate the integrals using following properties of Hermite polynomials
\begin{align}
  \int_{-\infty}^{\infty}
  w(\xi) \He_k(\xi)\He_l(\xi) d\xi = \sqrt{2\pi}k!\delta_{kl}, 
\end{align}
and
\begin{align}
  &\int_{-\infty}^{\infty}
  w(\xi) \He_m(\xi)\He_k(\xi)\He_l(\xi) d\xi\notag\\& =
  \frac{\sqrt{2\pi}m!k!l!}{(j-m)!(j-k)!(j-l)!}, 
\end{align}
where $j = (m+k+l)/2$. We then obtain
% \begin{align}
%  \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
%  \left(1+\hat{k}^2\right) = \delta q h \hat{k} \frac{n!}{(n-1)!}.
% \end{align
 \begin{align}
  \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
  \left(1+\hat{k}^2\right) = \delta q h \hat{k} n.
 \end{align}
Inserting the original eigenfrequency $\sigma = \pm
\sqrt{n}\Omega_k(1+\hat{k}^2)^{-1/2}$ gives
\begin{align}
  \frac{\imag(\delta\sigma)}{\Omega_k} = \pm \frac{1}{2}\delta q h
  \sqrt{n} \frac{\hat{k}}{\sqrt{1+\hat{k}^2}}. 
\end{align}
This result agrees with Eq. \ref{simple_growth}.  
%for $n=1$, since the
%function $\He_1$ solves the governing equation exactly. 
For $\hat{k}\gg 1$ we have
\begin{align}
  \frac{\imag(\delta\sigma)}{\Omega_k} \simeq \pm \frac{1}{2}\delta q h
  \sqrt{n} \sgn{\hat{k}}. 
\end{align}

### 2014-11-30 setup.tex
extensions

 \section{Baroclinic disk equilibra}
 We consider power-law disks with midplane density
 \begin{align}
   \rho_0(r) = \rho_{00}\left(\frac{r}{r_0}\right)^p,
 \end{align}
 where $\rho_{00}$ is the midplane density at the fiducial radius
 $r=r_0$, and midplane sound-speed 
 \begin{align}
   c_{s0}^2(r)=c_{s00}^2\left(\frac{r}{r_0}\right)^q, 
 \end{align}
 where $c_{s00}$ is the midplane sound-speed at $r=r_0$. 
 The three-dimensional equilibrium pressure $P$ and density $p$ are
 related by 
 \begin{align}
   P(r,z) = 
   \frac{c_{s0}^2(r)}{\Gamma}\rho_0(r)^{1-\Gamma}\rho^\Gamma(r,z), 
 \end{align}
 where we refer to $\Gamma$ as the polytropic index.

 \subsection{Locally isothermal disks}
 For the special case $\Gamma=1$ the disk is isothermal in the
 vertical direction. The equilibrium density $\rho(r,z)$ and rotation
 $\Omega(r,z)$ is then 
 \begin{align}
   \rho(r,z) &=
   \rho_0(r)\exp{\left[\frac{v_k^2}{c_{s0}^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]}\\   
   \Omega^2(r,z)& = \Omega_k^2(r) \left[(p+q)h^2(r) + (1 + q) -
     \frac{1}{\sqrt{1+z^2/r^2}}\right],
 \end{align}
 where $\Omega_k\equiv\sqrt{GM/r^3}=v_k/r$ is the Keplerian rotation 
 and 
 \begin{align}
   h(r) = \frac{c_{s0}(r)}{v_k(r)} 
 \end{align}
 is the local disk aspect ratio. The disk scale height is $H=hr$.

 The radial shear is 
 \begin{align}
   S(r,z) \equiv r\frac{\p\Omega}{\p r},  
 \end{align}
 and we note that the square of the epicyclic frequency $\kappa$ is
 given by
 \begin{align}
   \kappa^2(r,z) \equiv 2\Omega(2\Omega + S). 
 \end{align}
 The vertical shear is 
 \begin{align}
   \frac{\p\Omega^2}{\p z} =
   \frac{q\Omega_k^2z}{r^2\left(1+z^2/r^2\right)^{3/2}}. 
 \end{align}

 \section{The vertically global shearing box} 
% We are interested in the stability problem on radial length-scales
% much smaller than the local disk scale-height $H$. 
 We work within the framework of the \emph{vertically global shearing
   box} (VGSB) developed by \cite{mcnally14}. 
 In this approximation the background density, pressure, radial shear
 and vertical shear are all assumed to only 
 depend on height. That is, the expressions for these quantities are
 evaluated at $r=r_0$.

 The axisymmetric VGSB equations read
 \begin{align}
   & \frac{\p\rho}{\p t} + \frac{\p}{\p x}\left(\rho v_x\right) +
   \frac{\p}{\p z}\left(\rho v_z\right)=0,\\
   & \frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} +
   v_zr_0\frac{d\Omega_0}{dz} \hat{\bm{y}} = - 
   2\Omega_0\hat{\bm{z}}\times\bm{v} - S_0v_x\hat{\bm{y}} \notag\\ 
   &\phantom{\frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} +
   v_zr_0\frac{d\Omega_0}{dz} \hat{\bm{y}}=}-\frac{1}{\rho}\nabla P - \frac{\p\Phi_0}{\p z}\hat{\bm{z}},\\
   & \frac{\p P}{\p t} + \bm{v}\cdot\nabla P = -\gamma P
   \nabla\cdot\bm{v} -
   \frac{\rho}{t_c}\left(\frac{P}{\rho}-\left.\frac{P}{\rho}\right|_{t=0}\right), 
 \end{align}
 where $\bm{v}$ is the velocity field relative to the bulk flow
 (i.e. $\bm{v}=\bm{0}$ in equilibrium),  
 subscript $0$ indicates evaluation at $r=r_0$. In the energy equation
 the ratio of specific heats $\gamma$ is assumed constant, 
 and we have included a thermal relaxation term with timescale $t_c$.

@@

 {\bf below: at the moment only valid for locally isothermal
   equilibrium ($\Gamma=1$)} 
 Writing $\delta P /\rho \equiv W$ and $c_s^2\delta\rho/\rho\equiv Q$, 
 the linearized system of equations is 
 \begin{align}
   -&\ii\sigma \frac{Q}{c_s^2} + \ii k_x \delta v_x + \frac{d\dd
     v_z}{dz} + \frac{d\ln{\rho}}{dz}\delta v_z = 0,\\
   &\ii\sigma \delta v_x + 2\Omega\delta v_y = \ii k_x W,\\
   -&\frac{\kappa^2}{2\Omega}\delta v_x + \ii \sigma\delta v_y =
   r\frac{d\Omega}{dz}\delta v_z, \\
   & \ii\sigma\delta v_z = \frac{dW}{dz} +
   \frac{d\ln{\rho}}{dz}\left(W-Q\right), \\
   -&\ii\sigma W + \frac{1}{t_c}\left(W-Q\right) = -\ii\sigma\gamma Q +
   c_s^2\frac{d\ln{\rho}}{dz}\left(\gamma-1\right)\delta v_z.
 \end{align}

 It is convenient to express $\delta v_x$ as a function of $W$ and
 $\delta v_z$,
 \begin{align}
   \delta v_x = -\frac{1}{D}\left(\sigma k_x W +
     r\frac{d\Omega^2}{dz}\delta v_z\right),
 \end{align}
 where
 \begin{align}
   D \equiv \kappa^2 - \sigma^2.
 \end{align}

>>>>

 \section{Baroclinic disk equilibra}
 We consider power-law disks with midplane density
 \begin{align}
   \rho_0(r) = \rho_{00}\left(\frac{r}{r_0}\right)^p,
 \end{align}
 where $\rho_{00}$ is the midplane density at the fiducial radius
 $r=r_0$, and midplane sound-speed 
 \begin{align}
   c_{s0}^2(r)=c_{s00}^2\left(\frac{r}{r_0}\right)^q, 
 \end{align}
 where $c_{s00}$ is the midplane sound-speed at $r=r_0$. We also
 define
\begin{align}
  \epsilon(r) \equiv \frac{c_{s0}(r)}{v_k(r)}
\end{align}
to parametrize the magnitude of the sound speed, where
$v_k=r\Omega_k=\sqrt{GM/r}$ is the Keplerian azimuthal velocity.

The three-dimensional equilibrium pressure $P$ and density $p$ are
related by 
\begin{align}
  P(r,z) = 
  \frac{c_{s0}^2(r)}{\Gamma}\rho_0(r)^{1-\Gamma}\rho^\Gamma(r,z), 
\end{align}
where we refer to $\Gamma$ as the polytropic index. We define the
equilibrium sound-speed
\begin{align}
  c_s^2\equiv \frac{dP}{d\rho} =
  c_{s0}^2(r)\left(\frac{\rho}{\rho_0}\right)^{\Gamma-1}.  
\end{align}

The equilibrium density is given by vertical hydrostatic balance. This
gives, for $\Gamma\neq1$,  
\begin{align}\label{eqm_dens}
  &\left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} = 1 +
  \frac{\left(\Gamma-1\right)}{\epsilon^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right). 
\end{align}
Note that for $\Gamma\neq1$ the disk has a surface $z=H(r)$ such that
$\rho(r,H)=0$. If we parametrize
\begin{align}
  H(r)\equiv h r,
\end{align}
then the imposed sound-speed and disk thickness are related by 
\begin{align}
  h^2(r) = \left[1-\frac{\epsilon^2(r)}{\Gamma-1}\right]^{-2}-1. 
\end{align}

The equilibrium rotation profile is obtained by balancing radial
pressure force, centrifugal force and radial gravity from the central
star. This gives 
\begin{align}\label{eqm_rot}
  \frac{\Omega^2(r,z)}{\Omega_k^2(r)}=1 +
  \left(p+\frac{s}{\Gamma}\right)\epsilon^2(r) 
  +\frac{s}{\Gamma} \left(1-\frac{1}{\sqrt{1+z^2/r^2}}\right), 
\end{align}
where $s\equiv q+p(1-\Gamma)$. We also define the radial shear rate  
\begin{align}
  S(r,z) \equiv r\frac{\p\Omega}{\p r},  
\end{align}
and the square of the epicyclic frequency $\kappa$ 
\begin{align}
  \kappa^2(r,z) \equiv 2\Omega(2\Omega + S). 
\end{align}
We note that rotation profile generally depends on both $r$ and $z$,
unless the equilibrium is barotropic with $s=0$, in which case
$\Omega$ only depends on $r$.

\subsection{Vertically isothermal disks}
For the special case $\Gamma=1$ the disk is isothermal in the
vertical direction. The equilibrium density $\rho(r,z)$ and rotation
 $\Omega(r,z)$ is then 
 \begin{align}
   \frac{\rho(r,z)}{\rho_0(r)} &= \exp{\left[\frac{1}{\epsilon^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]}\\   
   \frac{\Omega^2(r,z)}{\Omega_k^2(r)}& =1+ (p+q)\epsilon^2(r) + q\left(1 -
     \frac{1}{\sqrt{1+z^2/r^2}}\right),
 \end{align}
which may be obtained from Eq. \ref{eqm_dens} and Eq. \ref{eqm_rot} by
taking the limit $\Gamma\to 1$.

The vertically isothermal disk has no surface, but we can define a
characteristic scale-height
\begin{align}
  H \equiv \epsilon r \quad\quad(\Gamma \equiv 1),   
\end{align} 
and in this case $h = \epsilon$.

% where $\Omega_k\equiv\sqrt{GM/r^3}=v_k/r$ is the Keplerian rotation 
 % and 
 % \begin{align}
 %   h(r) = \frac{c_{s0}(r)}{v_k(r)} 
 % \end{align}
 % is the local disk aspect ratio. 
%The disk scale height is $H=hr$.

 % The vertical shear is 
 % \begin{align}
 %   \frac{\p\Omega^2}{\p z} =
 %   \frac{q\Omega_k^2z}{r^2\left(1+z^2/r^2\right)^{3/2}}. 
 % \end{align}

 \section{The vertically global shearing box} 
% We are interested in the stability problem on radial length-scales
% much smaller than the local disk scale-height $H$. 
 We work within the framework of the \emph{vertically global shearing
   box} (VGSB) developed by \cite{mcnally14}. In this approximation
 one sets up a local Cartesian co-ordinate  system $(x,y,z)$ centered
 on a fiducial point $(r,\phi,z)=(r_0,\phi_0,0)$ in the global disk, with $\phi_0(t) =
 \Omega(r_0,0)t$. The local $x,y,z$ axis correspond to the
 global radial, azimuthal and vertical directions.

 In this shearing box the equilibrium density, pressure and 
 temperatures are assumed to depend only on $z$, by evaluating these
 quantities at $r=r_0$. The background rotation is approximated by a
 linear shear in $x$, but left unaltered in $z$. 
% This corresponds to
%  setting 
% \begin{align}
%   \rho(r,z)\to\rho(r_0,z),\,P \to P(r_0,z), c_s(r,z)^2 \to
%   c_s^2(r_0,z),  \\
% \end{align}
 The axisymmetric VGSB equations read
 \begin{align}
   & \frac{\p\rho}{\p t} + \frac{\p}{\p x}\left(\rho v_x\right) +
   \frac{\p}{\p z}\left(\rho v_z\right)=0,\\
   & \frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} +
   v_zr_0\frac{d\Omega_0}{dz} \hat{\bm{y}} = - 
   2\Omega_0\hat{\bm{z}}\times\bm{v} - S_0v_x\hat{\bm{y}} \notag\\ 
   &\phantom{\frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} +
   v_zr_0\frac{d\Omega_0}{dz} \hat{\bm{y}}=}-\frac{1}{\rho}\nabla P - \frac{\p\Phi_0}{\p z}\hat{\bm{z}},\\
   & \frac{\p P}{\p t} + \bm{v}\cdot\nabla P = -\gamma P
   \nabla\cdot\bm{v} -
   \frac{\rho}{t_c}\left(\frac{P}{\rho}-\left.\frac{P}{\rho}\right|_{t=0}\right), 
 \end{align}
 where $\bm{v}$ is the velocity field relative to the bulk flow
 (i.e. $\bm{v}=\bm{0}$ in equilibrium),  
 subscript $0$ indicates evaluation at $r=r_0$. In the energy equation
 the ratio of specific heats $\gamma$ is assumed constant, 
 and we have included a thermal relaxation term with timescale $t_c$.

@@

 % {\bf below: at the moment only valid for locally isothermal
 %  equilibrium ($\Gamma=1$)} 
 Writing $\delta P /\rho \equiv W$ and $c_s^2\delta\rho/\rho\equiv Q$, 
 the linearized system of equations is 
 \begin{align}
   -&\ii\sigma \frac{Q}{c_s^2} + \ii k_x \delta v_x + \frac{d\dd
     v_z}{dz} + \frac{d\ln{\rho}}{dz}\delta v_z = 0,\\
   &\ii\sigma \delta v_x + 2\Omega\delta v_y = \ii k_x W,\\
   -&\frac{\kappa^2}{2\Omega}\delta v_x + \ii \sigma\delta v_y =
   r\frac{d\Omega}{dz}\delta v_z, \\
   & \ii\sigma\delta v_z = \frac{dW}{dz} +
   \frac{d\ln{\rho}}{dz}\left(W-Q\right), \\
   -&\ii\sigma W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\sigma\frac{\gamma}{\Gamma} Q +
   c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma}-1\right)\delta v_z.
 \end{align}
It is understood that the quantities $\rho,\,c_s^2,\,\Omega,\kappa$
are evaluated at $r=r_0$ using their global expressions above.

We can eliminate the perturbed velocities between the linearized
equations to obtain a pair or ordinary differential equations
for $W,Q$ as
\begin{align}
  &\frac{\sigma^2}{c_s^2}Q + \frac{\sigma^2k_x^2}{D}W \notag\\ 
  &=\left(\frac{\ii
    k_x r}{D}\frac{d\Omega^2}{dz} -
  \frac{d\ln{\rho}}{dz}\right)\left[\frac{dW}{dz} +
  \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right] \notag\\
&- \frac{d^2W}{dz^2} - \frac{d\ln{\rho}}{dz}\left(\frac{dW}{dz} -
  \frac{dQ}{dz}\right) - \frac{d^2\ln{\rho}}{dz^2}\left(W-Q\right),\label{ode_w}\\
&\sigma^2W - \frac{\gamma}{\Gamma}\sigma^2 Q +
\frac{\ii\sigma}{t_c}\left(W-\frac{Q}{\Gamma}\right)\notag\\
&=c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma} - 1\right) 
\left[\frac{dW}{dz} + \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right],\label{ode_Q} 
\end{align}
 % It is convenient to express $\delta v_x$ as a function of $W$ and
 % $\delta v_z$,
 % \begin{align}
 %   \delta v_x = -\frac{1}{D}\left(\sigma k_x W +
 %     r\frac{d\Omega^2}{dz}\delta v_z\right),
 % \end{align}
where
\begin{align}
  D \equiv \kappa^2 - \sigma^2.
\end{align}
